\double{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}{Автоматический~анализ~слабых~моделей памяти~параллельного~программирования}

### 2018-06-18 slides/slides.tex
Finish slides.

@@

\usepackage{colortbl}

\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
\definecolor{ltblue}{RGB}{101,147,245}
\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
\definecolor{dkred} {rgb} {0.6, 0.1, 0}
\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}

@@

\usepackage{fancyvrb}%fancy verbatim
%\makeatletter
%\newcommand{\verbatimfont}[1]{\renewcommand{\verbatim@font}{\ttfamily\scriptsize#1}}
%\makeatother
% redefine \VerbatimInput
\RecustomVerbatimCommand{\VerbatimInput}{VerbatimInput}{fontsize=\scriptsize}

\usepackage{pmboxdraw}
\usepackage{newunicodechar}
\newunicodechar{└}{\textSFii}
\newunicodechar{├}{\textSFviii}
\newunicodechar{─}{\textSFx}

@@

\usepackage[backend=biber]{biblatex}
\addbibresource{../bib/ms-thesis.bib}

@@

%\titlegraphic{\includegraphics[height=1cm,width=2cm]{logo1}}
%\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}
\title{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}

% A subtitle is optional and this may be deleted
%\subtitle{Optional Subtitle}
\author{\textbf{Artem Yushkovskiy}\inst{1,2} \\ 
{\scriptsize MSc Candidate}
\\ \vspace{1em}
{\footnotesize\raggedleft Supervisors: \textbf{Assoc. Prof. Keijo Heljanko}\inst{1} \newline
\hphantom{Supervis} \textbf{Docent Igor I. Komarov}\inst{2} }
}%\small
%{\footnotesize%\raggedleft%does not work
%Supervisor 1: Assoc. Prof. Keijo Heljanko\inst{1} \newline
%Supervisor 2: Docent Igor I. Komarov\inst{2}
%}

\institute
{
  \inst{1}%
  \textbf{Aalto University} (Espoo, Finland) \\
  Department of Computer Science, \\
  School of Science
  \and
  \inst{2}%
  \textbf{ITMO University} (Saint Petersburg, Russia) \\
  Faculty of Information Security \\
  and Computer Technologies
}

\date{\scriptsize Espoo, Saint Petersburg, 2018}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

@@

%\begin{frame}{Outline}
%  \tableofcontents
%\end{frame}

%\section{Introduction}
%\subsection{The problem}

\begin{frame}{Task statement \rus{(Цель работы)}}
\begin{itemize}
\item Improve the proof-of-concept memory model-aware analysis tool \texttt{Porthos}~\cite{Porthos17a} by extending support for the input language.
\end{itemize}
\end{frame}

\begin{frame}{Task specification \rus{(Задачи работы)}}
\begin{itemize}
\item Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};
\item Review existing tools for memory model-aware analysis;
%\item Examine the architecture of \texttt{Porthos}, its strengths and weaknesses;
%\item Design a new architecture for \texttt{PorthosC} that allow to extend the input language to the (large subset of) C language, be robust, transparent, efficient and extensible.
\item Develop a \textit{C compiler infrastructure} as a part of an abstract interpretation engine for the new tool \tool{PorthosC};
\item Improve the SMT-encoding scheme for an \textit{arbitrary control-flow};
\item Enhance performance, extensibility, reliability and maintainability of the tool.
\end{itemize}
\end{frame}

@@

\begin{frame}{Verification of concurrent software} {Motivating example: Write-write reordering (compiler  relaxations)}
%1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
%2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/
\begin{table}
\centering
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
\tikzmark{p0}$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\tikzmark{q0} \\
\tikzmark{p1}$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\tikzmark{q1} \\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
\end{table}

\onslide<2->
\begin{minipage}{.25\textwidth}
\small
\begin{table}
\only<1>{\caption*{~\\~}}
\only<2>{\caption*{Single thread\\(no concurrency)}}
\only<3>{\caption*{Sequential Consistency\\(classical concurrency)}}
\only<4->{\caption*{~\\~}}
\only<4->{
  \begin{tikzpicture}[overlay, remember picture]
    \draw [<->,thick,red,looseness=3] ([xshift=-5pt,yshift=3pt]{pic cs:p0}) to[bend right=90] ([xshift=-5pt]{pic cs:p1});
    \draw [<->,thick,red,looseness=3] ([xshift=8pt,yshift=3pt]{pic cs:q0}) to[bend left=90] ([xshift=5pt]{pic cs:q1});
  \end{tikzpicture}%
}
%"A multiprocessor is sequentially consistent if the result of any
%execution is the same as some sequential order, and within any
%processor, the operations are executed in program order"
%– Lamport, 1979
\begin{tabular}{ l r }
$\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$
\onslide<3->
\\
$\b{p_0}, \r{q_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\b{p_0}, \r{q_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
\end{tabular}
\end{table}
\end{minipage}
%
\onslide<4->
\begin{minipage}{.6\textwidth}
\small
%"delay the effect of a store past any load from a different location."
\begin{table}
%http://events17.linuxfoundation.org/sites/events/files/slides/dbueso-elc2016-membarriers-final.pdf
%SPARC, x86 (Intel, AMD)
%Similar to SC, but:
%‒ Loads may be reordered with writes.
%Relaxed Models
% Arbitrary reorder limited only by explicit memorybarrier
%instructions.
\caption*{Total Store Order\\(e.g., x86)}
\begin{tabular}{ | l r | l r | l r } % >{\columncolor{ltgrey}}
$\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}}  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
\end{tabular}
\end{table}
\end{minipage}

@@

\begin{frame}{Verification of concurrent software} {Motivating example: Store buffering (hardware relaxations)}
\begin{center}
\scalebox{0.75}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
}
\end{center}

\begin{figure}
\includegraphics[scale=0.37]{img/x86-arch-full.png}
\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}
\only<2->{
    \begin{tikzpicture}[overlay]
    \node at (0.1, 3.1) {\fcolorbox{yellow}{yellow}{
        \includegraphics[scale=0.2]{img/table.png}} };
    \end{tikzpicture}
}
\end{figure}

@@

\begin{frame}{Weak memory model-aware analysis} {Event-based program representation}

\begin{itemize}
  \addtolength{\tabcolsep}{0pt}
  \item \textbf{Event} $\in \mathbb{E}$, a low-level primitive operation:
    \begin{itemize}
      \item \textit{memory event} $\in \mathbb{M} = \mathbb{R} \cup \mathbb{W}$: access to a local/shared memory, \\
      \item \textit{computational event} $\in \mathbb{C}$: computation over local memory, and \\
      \item \textit{barrier event} $\in \mathbb{B}$: synchronisation fences;
     \end{itemize}
  \item \textbf{Relation} $\subseteq \mathbb{E} \times \mathbb{E}$: 
    \begin{itemize}
    \item \textit{basic relations}:%, computed from the program model:
      \begin{itemize}
        \item \textit{program-order} relation $\texttt{po} \subseteq \mathbb{E} \times \mathbb{E}$: (control-flow),
        \item \textit{read-from} relation $\texttt{rf} \subseteq \mathbb{W} \times \mathbb{R}$: (data-flow), and
        \item \textit{coherence-order} relation $\texttt{co} \subseteq \mathbb{W} \times \mathbb{W}$: (data-flow);
      \end{itemize}
    \item \textit{derived relations}:%, computed from the basic relations using the following operators:
      \begin{itemize}
        \item \textit{union} \texttt{r1\,|\,r2},
        \item \textit{sequence} \texttt{r1} ; \texttt{r2},
        %\item \textit{inverse} \texttt{r}\^{\texttt{-1}},
        \item \textit{transitive closure} \texttt{r+},
        \item $\cdots$;
      \end{itemize}
    \end{itemize}
  %\item \textbf{Candidate execution}, a sequence of guesses of relations;
  \item \textbf{Assertion} over relations or sets of events: 
    \begin{itemize}
    \item \textit{acyclicity}, \textit{irreflexivity} or \textit{emptiness} 
    \end{itemize}
\end{itemize}

@@

\begin{frame}{Weak memory model-aware analysis} {Testing candidate executions}
\vspace{10pt}
\begin{minipage}{.33\linewidth}
\hfill
\end{minipage}
%
\noindent\begin{minipage}{.39\linewidth}
\scalebox{0.65}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
}
\end{minipage}
%
\noindent\begin{minipage}{.23\linewidth}
\only<2->{
\scriptsize
SC model: \\
... \\
\texttt{fr = (rf\^\,-1; co)} \\
\color{red}\texttt{acyclic(fr $\cup$ po)}
}
\end{minipage}
\vspace{-5pt}
\only<1>{
\begin{figure}
%\centering
\includegraphics[width=\linewidth]{img/candidate.png}
\caption{The four candidate executions allowed under x86-TSO}
\end{figure}
}
\only<2->{
\begin{figure}
%\centering
\includegraphics[width=\linewidth]{img/candidate-red.png}
\caption{The four candidate executions allowed under x86-TSO}
\end{figure}
}
\only<3>{
  \begin{tikzpicture}[overlay]
  \node at (6.1, 4.1) {\fcolorbox{yellow}{yellow}{
    \includegraphics[scale=0.11]{img/non-portable.png}} };
  \end{tikzpicture}
}

@@

\begin{frame}{Portability analysis} {The \tool{Porthos} tool}
\begin{minipage}{.49\textwidth}
\only<1->{
  \begin{figure}
  \includegraphics[height=0.4\textheight,keepaspectratio]{img/the-problem.png}
  \caption{The illustration of the portability problem~\cite{Porthos17slides}}
  \end{figure}
}
\end{minipage}
%
\begin{minipage}{.5\textwidth}
\only<2->{
  \begin{itemize}
    \begin{definition}[Portability~\cite{Porthos17a}]
    %SAY: Let $\textit{cons}_{\mathcal{M}}(P)$ be a function that calculates the set of executions of the program $P$ that are consistent under the memory model $\mathcal{M}$
    %Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
    The program $P$ is portable from the source memory model $\mathcal{M_S}$ to the target one $\mathcal{M_T}$ if
    $\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
    \end{definition}
    \item Portability as an SMT-based bounded reachability problem: \\
    $\phi = $\only<3>\yellowemph{$\phi_{CF}$}$\land $\only<3>\yellowemph{$\phi_{DF}$}$\land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$
    \item $\texttt{SAT}(\phi) \Longrightarrow $ portability bug
  \end{itemize}
}
\end{minipage}
\end{frame}

>>>>

%\def\IsRus{1}
\newcommand{\double}[2]{\ifdefined\IsRus{#2}\else{#1}\fi}

@@

\usepackage{colortbl}
\definecolor{ltgrey}{rgb}{0.85,0.85,0.85}
\definecolor{ltblue}{RGB}{101,147,245}
\definecolor{dkblue}{rgb} {0,   0.1, 0.6}
\definecolor{dkred} {rgb} {0.6, 0.1, 0}
\renewcommand{\r}[1]{\ensuremath{\textcolor{dkred}{#1}}}
\renewcommand{\b}[1]{\ensuremath{\textcolor{dkblue}{#1}}}

@@

@@

\usepackage[backend=biber]{biblatex}
\addbibresource{../bib/ms-thesis.bib}
\renewcommand*{\bibfont}{\scriptsize}

@@

%\titlegraphic{\includegraphics[height=1cm,width=2cm]{logo1}}
%\titlegraphicii{\includegraphics[height=1cm,width=2cm]{logo2}}
\title{\double{\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}}{Автоматический~анализ~слабых~моделей памяти~параллельного~программирования}}

% A subtitle is optional and this may be deleted
%\subtitle{Optional Subtitle}
\author{
\double{
\textbf{Artem Yushkovskiy}\inst{1,2} \\ 
{\scriptsize MSc Candidate}
\\ \vspace{1em}
{\footnotesize\raggedleft Supervisors: \textbf{Assoc. Prof. Keijo Heljanko}\inst{1} \newline
\hphantom{Supervis} \textbf{Docent Igor I. Komarov}\inst{2} }
}
{\textbf{Юшковский А. В.}\inst{1,2} \\ 
{\scriptsize магистрант}
\\ \vspace{1em}
{\footnotesize\raggedleft Руководители: \textbf{проф. Кеийо Хельянко}\inst{1} \newline
\hphantom{Руководители} \textbf{доц. Комаров И. И.}\inst{2} }
}
}

\institute
{
  \inst{1}%
  \double{\textbf{Aalto University} (Espoo, Finland) \\
    Department of Computer Science, \\
    School of Science
  }
  {\textbf{Университет Аалто} (Эспоо, Финляндия) \\
    Факультет Информатики, Школа Наук
  }
  \and
  \inst{2}%
  \double{\textbf{ITMO University} (Saint Petersburg, Russia) \\
    Faculty of Information Security \\
    and Computer Technologies
  }
  {\textbf{Университет ИТМО} (Санкт-Петебург, Россия) \\
    Факультет Безопасности Информационных Технологий
  }
}

\date{\scriptsize \double{Espoo, Saint Petersburg\\2018}{Эспоо, Санкт Петербург\\2018}}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

@@

\begin{frame}{\double{Task statement}{Цель работы}}
\begin{itemize}
\item \double{Improve the proof-of-concept memory model-aware analysis tool \tool{Porthos}~\cite{Porthos17a} by extending support for the input language.}
{Усовершенствовать прототип статического анализатора кода  \tool{Porthos}~\cite{Porthos17a}, учитывающий при анализе модель памяти вычислительной среды, путем расширения поддержки входного языка.}
\end{itemize}
\end{frame}

\begin{frame}{\double{Task specification}{Задачи работы}}
\begin{itemize}
\item \double{Study the general framework for memory model-aware analysis of concurrent programs~\cite{alglave2010shared};}
{Изучить существующие подходы к автоматическому проведению анализа кода параллельных программ с учетом модели памяти вычислительной среды~\cite{alglave2010shared};}
\item \double{Review existing tools for memory model-aware analysis;}
{Исследовать существующие инструменты анализа кода с учетом модели памяти;}
%\item Examine the architecture of \texttt{Porthos}, its strengths and weaknesses;
%\item Design a new architecture for \texttt{PorthosC} that allow to extend the input language to the (large subset of) C language, be robust, transparent, efficient and extensible.
\item \double{Develop a \textit{C compiler infrastructure} as a part of an abstract interpretation engine for the new tool \tool{PorthosC};}
{Разработать инфраструктуру C-компилятора как часть модуля абстрактной интерпретации нового анализатора \tool{PorthosC};}
\item \double{Improve the SMT-encoding scheme for an \textit{arbitrary control-flow};}
{Адаптировать схему кодирования произвольного графа потока управления в SMT-формулу;}
\item \double{Enhance performance, extensibility, reliability and maintainability of the tool.}
{Сохранить и улучшить функциональные характеристики анализатора (производительность, расширяемость, надежность и поддерживаемость).}
\end{itemize}
\end{frame}

@@

\begin{frame}{\double{Verification of concurrent software}{Верификация параллельных программ}}{\double{Motivating example: Write-write reordering (compiler  relaxations)}{Пример: Перестановка типа запись-запись (оптимизация на уровне компилятора)}}
%1) mckenney  http://www.rdrop.com/users/paulmck/scalability/paper/whymb.2010.06.07c.pdf
%2) http://preshing.com/20120515/memory-reordering-caught-in-the-act/
\begin{table}
\centering
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
\tikzmark{p0}$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\tikzmark{q0} \\
\tikzmark{p1}$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\tikzmark{q1} \\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
\end{table}

\onslide<2->
\begin{minipage}{.25\textwidth}
\small
\begin{table}
\only<1>{\caption*{~\\~}}
\only<2>{\double{\caption*{Single thread\\(no concurrency)}}{\caption*{Один поток\\(отсутствие параллелизма)}}}
\only<3>{\double{\caption*{Sequential Consistency\\(classical concurrency)}}{\caption*{Последовательная Согласованность\\(классический параллелизм)}}}
\only<4->{\caption*{~\\~}}
\only<4->{
  \begin{tikzpicture}[overlay, remember picture]
    \draw [<->,thick,red,looseness=3] ([xshift=-5pt,yshift=3pt]{pic cs:p0}) to[bend right=90] ([xshift=-5pt]{pic cs:p1});
    \draw [<->,thick,red,looseness=3] ([xshift=8pt,yshift=3pt]{pic cs:q0}) to[bend left=90] ([xshift=5pt]{pic cs:q1});
  \end{tikzpicture}%
}
%"A multiprocessor is sequentially consistent if the result of any
%execution is the same as some sequential order, and within any
%processor, the operations are executed in program order"
%– Lamport, 1979
\begin{tabular}{ l r }
$\b{p_0}, \b{p_1}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$
\onslide<3->
\\
$\b{p_0}, \r{q_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\b{p_0}, \r{q_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \b{p_1}, \r{q_1}$ & $(\b{1}; \r{1})$ \\
$\r{q_0}, \b{p_0}, \r{q_1}, \b{p_1}$ & $(\b{1}; \r{1})$ \\
\end{tabular}
\end{table}
\end{minipage}
%
\onslide<4->
\begin{minipage}{.6\textwidth}
\small
%"delay the effect of a store past any load from a different location."
\begin{table}
%http://events17.linuxfoundation.org/sites/events/files/slides/dbueso-elc2016-membarriers-final.pdf
%SPARC, x86 (Intel, AMD)
%Similar to SC, but:
%‒ Loads may be reordered with writes.
%Relaxed Models
% Arbitrary reorder limited only by explicit memorybarrier
%instructions.
\double{\caption*{Total Store Order\\(e.g., x86)}}{\caption*{Линейный порядок записи\\(напр., x86)}}
\begin{tabular}{ | l r | l r | l r } % >{\columncolor{ltgrey}}
$\underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_0}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \b{p_1}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_1}}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$ \\
$\r{q_0}, \r{q_1}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_0}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_1}}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$ \\
$\underline{\b{p_1}}, \r{q_0}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \b{p_1}, \underline{\r{q_0}}$ & $(\b{0}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\underline{\b{p_1}}, \r{q_0}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{0}; \r{1})$  &  $\b{p_0}, \underline{\r{q_1}}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{1})$  &  $\underline{\b{p_1}}, \underline{\r{q_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\r{q_0}, \underline{\b{p_1}}, \underline{\b{p_0}}, \r{q_1}$ & $(\b{1}; \r{1})$  &  $\underline{\r{q_1}}, \b{p_0}, \b{p_1}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}}  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\b{p_0}}, \underline{\r{q_0}}$ & {\only<5->\yellowemph{$(\b{0}; \r{0})$}} \\
$\r{q_0}, \underline{\b{p_1}}, \r{q_1}, \underline{\b{p_0}}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \b{p_0}, \underline{\r{q_0}}, \b{p_1}$ & $(\b{1}; \r{0})$  &  $\underline{\r{q_1}}, \underline{\b{p_1}}, \underline{\r{q_0}}, \underline{\b{p_0}}$ & {\only<5>\yellowemph{$(\b{0}; \r{0})$}} \\
\end{tabular}
\end{table}
\end{minipage}

@@

\begin{frame}{\double{Verification of concurrent software}{Верификация параллельных программ}} {\double{Motivating example: Store buffering (hardware relaxations)}{Пример: Буферизация записи (оптимизация на уровне процессора)}}
\begin{center}
\scalebox{0.75}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
}
\end{center}

\begin{figure}
\includegraphics[scale=0.37]{img/x86-arch-full.png}
\double{\caption{An x86-TSO abstract machine~\cite{sewell2010x86}}}{\caption{Модель абстрактной машины процессора x86~\cite{sewell2010x86}}}
\only<2->{
    \begin{tikzpicture}[overlay]
    \node at (0.1, 3.1) {\fcolorbox{yellow}{yellow}{
        \includegraphics[scale=0.2]{img/table.png}} };
    \end{tikzpicture}
}
\end{figure}

@@

\begin{frame}{\double{Weak memory model-aware analysis}{Статический анализ с учетом слабой модели памяти}} {\double{Event-based program representation}{Событийная модель кода}}

\begin{itemize}
  \addtolength{\tabcolsep}{0pt}
  \item \textbf{\double{Event}}{\textbf{Событие}} $\in \mathbb{E}$, \double{a low-level primitive operation}{низкоуровневая примитивная операция}:
    \begin{itemize}
      \item \textit{\double{memory event}{событие памяти}} $\in \mathbb{M} = \mathbb{R} \cup \mathbb{W}$: access to a local/shared memory, \\
      \item \textit{\double{computational event}{событие вычисления}} $\in \mathbb{C}$: \double{computation over local memory, and}{вычисление в локальной памяти} \\
      \item \textit{\double{barrier event}{событие синхронизации}} $\in \mathbb{B}$: \double{synchronisation fences}{барьеры синхронизации};
     \end{itemize}
  \item \textbf{\double{Relation}{Отношение}} $\subseteq \mathbb{E} \times \mathbb{E}$: 
    \begin{itemize}
    \item \textit{\double{basic relations}{базовые отношения}}:%, computed from the program model:
      \begin{itemize}
        \item \textit{program-order} \double{relation}{отношение} $\texttt{po} \subseteq \mathbb{E} \times \mathbb{E}$: (\double{control-flow}{поток управления}),
        \item \textit{read-from} \double{relation}{отношение} $\texttt{rf} \subseteq \mathbb{W} \times \mathbb{R}$: (\double{data-flow, and}{поток данных, и})
        \item \textit{coherence-order} \double{relation}{отношение} $\texttt{co} \subseteq \mathbb{W} \times \mathbb{W}$: (\double{data-flow}{поток данных});
      \end{itemize}
    \item \textit{\double{derived relations}{производные отношения}}:%, computed from the basic relations using the following operators:
      \begin{itemize}
        \item \textit{\double{union}{объединение}} \texttt{r1\,|\,r2},
        \item \textit{\double{sequence}{последовательность}} \texttt{r1} ; \texttt{r2},
        %\item \textit{inverse} \texttt{r}\^{\texttt{-1}},
        \item \textit{\double{transitive closure}{транзитивное замыкание}} \texttt{r+},
        \item $\cdots$;
      \end{itemize}
    \end{itemize}
  %\item \textbf{Candidate execution}, a sequence of guesses of relations;
  \item \double{\textbf{Assertion} over relations or sets of events}{\textbf{Ограничение} модели памяти, сформулированное в терминах отношений или множеств событий} : 
    \begin{itemize}
    \item \textit{\double{acyclicity}{ацикличность}}, \textit{\double{irreflexivity}{антирефлексивность}} or \textit{\double{emptiness}{отсутствие элементов во множестве}}.
    \end{itemize}
\end{itemize}

@@

\begin{frame}{\double{Weak memory model-aware analysis}{Статический анализ с учетом слабой модели памяти}} {\double{Testing candidate executions}{Тестирование путей выполнения программы}}
\vspace{10pt}
\begin{minipage}{.33\linewidth}
\hfill
\end{minipage}
%
\noindent\begin{minipage}{.39\linewidth}
\scalebox{0.65}{
\ttfamily
\begin{tabular}{ |>{\color{dkblue}}l | >{\color{dkred}}l| }
\hline
\multicolumn{2}{|c|}{ \{ x=0; y=0; \}} \tabularnewline \hline
P & Q \\ \hline
$p_0$: $x \leftarrow 1$   & $q_0$: $y \leftarrow 1$\\
$p_1$: $r_p \leftarrow y$ & $q_1$: $r_q \leftarrow x$\\
\hline
\multicolumn{2}{|c|}{ exists ($\b{r_p}=0 \land \r{r_q}=0$) } \\
\hline
\end{tabular}
}
\end{minipage}
%
\noindent\begin{minipage}{.23\linewidth}
\only<2->{
\scriptsize
SC \double{model}{модель}: \\
... \\
\texttt{fr = (rf\^\,-1; co)} \\
\color{red}\texttt{acyclic(fr $\cup$ po)}
}
\end{minipage}
\vspace{-5pt}
\only<1>{
\begin{figure}
%\centering
\includegraphics[width=\linewidth]{img/candidate.png}
\double{\caption{The four candidate executions allowed under x86-TSO}}{\caption{Четыре возможных выполнения программы, допустимых моделью x86-TSO}}
\end{figure}
}
\only<2->{
\begin{figure}
%\centering
\includegraphics[width=\linewidth]{img/candidate-red.png}
\double{\caption{The four candidate executions allowed under x86-TSO}}{\caption{Четыре возможных выполнения программы, допустимых моделью x86-TSO}}
\end{figure}
}
\only<3>{
  \begin{tikzpicture}[overlay]
  \node at (6.1, 4.1) {\fcolorbox{yellow}{yellow}{
    \includegraphics[scale=0.11]{img/non-portable.png}} };
  \end{tikzpicture}
}

@@

\begin{frame}{\double{Portability analysis}{Анализ портируемости}} {\double{The \tool{Porthos} tool}{Анализатор \tool{Porthos}}}
\begin{minipage}{.49\textwidth}
\only<1->{
  \begin{figure}
  \includegraphics[height=0.4\textheight,keepaspectratio]{img/the-problem.png}
  \caption{\double{The illustration of the portability problem~\cite{Porthos17slides}}{Иллюстрация проблемы определения портируемости программы~\cite{Porthos17slides}}}
  \end{figure}
}
\end{minipage}
%
\begin{minipage}{.5\textwidth}
\only<2->{
  \begin{itemize}
    \begin{definition}[\double{Portability}{Портируемость}~\cite{Porthos17a}]
    %SAY: Let $\textit{cons}_{\mathcal{M}}(P)$ be a function that calculates the set of executions of the program $P$ that are consistent under the memory model $\mathcal{M}$
    %Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models.
    \double{The program $P$ is portable from the source memory model $\mathcal{M_S}$ to the target model $\mathcal{M_T}$ if}
    {Программа $P$ называется портируемой с платформы $\mathcal{M_S}$ на платформу $\mathcal{M_T}$, если}
    $\textit{cons}_{\mathcal{M_T}}(P) \subseteq \textit{cons}_{\mathcal{M_S}}(P)$
    \end{definition}
    \item \double{Portability as an SMT-based bounded reachability problem}{Портируемость как проблема ограниченной достижимости на основе SMT}: \\
    $\phi = $\only<3>\yellowemph{$\phi_{CF}$}$\land $\only<3>\yellowemph{$\phi_{DF}$}$\land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$
    \item $\texttt{SAT}(\phi) \Longrightarrow $ \double{portability bug}{ошибка портирования}
  \end{itemize}
}
\end{minipage}
\end{frame}
